---
title: "Model Prediction + Feature Importance"
author: "shantinii"
format: html
editor: visual
---

load package

```{r}
library(caret)
library(e1071)
library(rpart)
library(randomForest)
library(xgboost)
library(pROC)
library(ggplot2)
library(gridExtra)
library(reshape2)
library(PRROC)
library(mice)
library(caret)
library(VIM)
library(missForest)
library(cowplot)
```

```{r}
#set working directory
setwd("../Myocardial Infraction/")
```

```{r}
options(warn = -1)
```

# Section 1: Data description

This study uses the “Myocardial Infarction Complications” dataset obtained from the UC Irvine Machine Learning Repository. The dataset contains 1701 patient records with 124 clinical features, including demographic information, ECG findings, laboratory results, underlying diseases and complications related to the myocardial infarction. The dataset also contains 15974 missing values, which represent a common challenge in real-world medical datasets (Golovenkin et al., 2020). Therefore, appropriate preprocessing and imputation methods are important to improve data quality before predictive analysis.

The objectives of this study are:

1\. To evaluate and determine the most suitable missing value imputation method for preprocessing the myocardial infarction dataset.\
2. To compare class imbalance handling techniques and determine the most effective method for improving predictive performance in an imbalanced myocardial infarction dataset.\
3. To develop and compare machine learning models for predicting mortality risk among acute myocardial infarction patients using clinical information available at the time of emergency room admission.\
4. To identify the most influential clinical predictors associated with mortality among acute myocardial infarction patients at the time of emergency room admission.

# Section 2: Imputation

LOAD DATA

```{r}
myocardial <- read.csv("myocardial_cleaned_100features.csv", header = TRUE, na.strings = c("?", "NA", ""))
print("Original dataset shape:")
print(dim(myocardial))
cat("Total missing cells before imputation:", sum(is.na(myocardial)), "\n\n")

```

## 2.1 Predictive Mean Matching (PMM)

```{r}
set.seed(42)

mice_pmm <- mice(myocardial, m = 5, maxit = 5, method = "pmm", printFlag = FALSE) 
imp_pmm <- complete(mice_pmm, 1)
```

## 2.2. Logistic Regression Chains (LogReg)

```{r}
# Setup workframe and drop ID variable 
logreg_df <- myocardial
logreg_df$ID <- NULL
```

```{r}
# Remove near zero variance columns using caret
nzv_cols <- nearZeroVar(logreg_df)
if(length(nzv_cols) > 0) {
  logreg_df <- logreg_df[, -nzv_cols]}
```

```{r}
# Find columns with exactly 2 values to set as factors for logreg
val_counts <- sapply(logreg_df, function(x) length(unique(na.omit(x))))
bin_cols   <- names(val_counts[val_counts == 2])
logreg_df[bin_cols] <- lapply(logreg_df[bin_cols], as.factor)
```

```{r}
# Initialize methods array
init_mice <- mice(logreg_df, maxit = 0)
mice_meth <- init_mice$method
mice_meth[bin_cols] <- "logreg"
```

```{r}
# Quick predictor matrix
pred_mat <- quickpred(logreg_df, mincor = 0.1, minpuc = 0.25)

set.seed(42)
mice_logreg <- mice(logreg_df, method = mice_meth, predictorMatrix = pred_mat, 
                    m = 5, maxit = 5, printFlag = FALSE)
imp_logreg  <- complete(mice_logreg)
```

```{r}
# Put back the ID column if it was removed
if ("ID" %in% colnames(myocardial)) {
  imp_logreg <- cbind(ID = myocardial$ID, imp_logreg)
}
```

## 2.3 & 2.4. Baseline Mean and Median Imputations

```{r}
imp_mean   <- myocardial
imp_median <- myocardial

num_cols <- sapply(myocardial, is.numeric)
```

```{r}
# Column mean replacement loop
imp_mean[num_cols] <- lapply(imp_mean[num_cols], function(x) {
  replace(x, is.na(x), mean(x, na.rm = TRUE))})
```

```{r}
# Column median replacement loop
imp_median[num_cols] <- lapply(imp_median[num_cols], function(x) {
  replace(x, is.na(x), median(x, na.rm = TRUE))})
```

## 2.5. K-Nearest Neighbors (KNN)

```{r}
set.seed(42)

# Run VIM kNN with k=5 
knn_raw <- kNN(myocardial, k = 5, imp_var = FALSE)
imp_knn <- knn_raw[, 1:ncol(myocardial)]
```

## 2.6. Random Forest (missForest)

```{r}
set.seed(42)

forest_run <- missForest(myocardial, verbose = FALSE)
imp_forest <- forest_run$ximp
```

## 2.7 Cleaning

```{r}
# Identify binary columns using the original dataset values
raw_counts <- sapply(myocardial, function(x) length(unique(na.omit(x))))
true_binary_features <- names(raw_counts[raw_counts == 2])
```

```{r}
# Simple loop function to fix data types, fractions, and trailing decimals
fix_binary_format <- function(df, bin_vars) {
  for (col in names(df)) {
    if (col %in% bin_vars) {
# Handle factors/characters safely, round any means/forest decimals, force integer type
      df[[col]] <- as.integer(round(as.numeric(as.character(df[[col]]))))
    }
  }
  return(df)
}
```

```{r}
# Apply format cleanup to all datasets before saving
imp_pmm    <- fix_binary_format(imp_pmm, true_binary_features)
imp_logreg <- fix_binary_format(imp_logreg, true_binary_features)
imp_mean   <- fix_binary_format(imp_mean, true_binary_features)
imp_median <- fix_binary_format(imp_median, true_binary_features)
imp_knn    <- fix_binary_format(imp_knn, true_binary_features)
imp_forest <- fix_binary_format(imp_forest, true_binary_features)
```

## 2.8 Visualisation

```{r}
plot_data <- data.frame(
  Original = myocardial$S_AD_ORIT,
  PMM = imp_pmm$S_AD_ORIT,
  LogReg = imp_logreg$S_AD_ORIT,
  Mean = imp_mean$S_AD_ORIT,
  Median = imp_median$S_AD_ORIT,
  KNN = imp_knn$S_AD_ORIT,
  missForest = imp_forest$S_AD_ORIT
)

p1 <- ggplot(plot_data, aes(x = Original))   + geom_histogram(fill = "#ad1538", color = "black", bins = 30) + ggtitle("Original") + theme_classic()
p2 <- ggplot(plot_data, aes(x = PMM))        + geom_histogram(fill = "#4a69bd", color = "black", bins = 30) + ggtitle("PMM Imputed")  + theme_classic()
p3 <- ggplot(plot_data, aes(x = LogReg))     + geom_histogram(fill = "#15ad4f", color = "black", bins = 30) + ggtitle("LogReg Imputed") + theme_classic()
p4 <- ggplot(plot_data, aes(x = Mean))       + geom_histogram(fill = "#1543ad", color = "black", bins = 30) + ggtitle("Mean Imputed")       + theme_classic()
p5 <- ggplot(plot_data, aes(x = Median))     + geom_histogram(fill = "#ad8415", color = "black", bins = 30) + ggtitle("Median Imputed")     + theme_classic()
p6 <- ggplot(plot_data, aes(x = KNN))        + geom_histogram(fill = "#ff69b4", color = "black", bins = 30) + ggtitle("KNN Imputed")        + theme_classic()
p7 <- ggplot(plot_data, aes(x = missForest)) + geom_histogram(fill = "#00ffff", color = "black", bins = 30) + ggtitle("missForest Imputed")   + theme_classic()

plot_grid(p1, p2, p3, p4, p5, p6, p7, nrow = 3, ncol = 3)
```

## 2.9 Statistical summary

```{r}
library(dplyr)

summary_stats <- data.frame(
  Method = c("Original", "PMM", "LogReg", "Mean", "Median", "KNN", "missForest"),
  N_Missing = c(
    sum(is.na(myocardial$S_AD_ORIT)),
    sum(is.na(imp_pmm$S_AD_ORIT)),
    sum(is.na(imp_logreg$S_AD_ORIT)),
    sum(is.na(imp_mean$S_AD_ORIT)),
    sum(is.na(imp_median$S_AD_ORIT)),
    sum(is.na(imp_knn$S_AD_ORIT)),
    sum(is.na(imp_forest$S_AD_ORIT))
  ),
  Mean = round(c(
    mean(myocardial$S_AD_ORIT, na.rm = TRUE),
    mean(imp_pmm$S_AD_ORIT),
    mean(imp_logreg$S_AD_ORIT),
    mean(imp_mean$S_AD_ORIT),
    mean(imp_median$S_AD_ORIT),
    mean(imp_knn$S_AD_ORIT),
    mean(imp_forest$S_AD_ORIT)
  ), 3),
  SD = round(c(
    sd(myocardial$S_AD_ORIT, na.rm = TRUE),
    sd(imp_pmm$S_AD_ORIT),
    sd(imp_logreg$S_AD_ORIT),
    sd(imp_mean$S_AD_ORIT),
    sd(imp_median$S_AD_ORIT),
    sd(imp_knn$S_AD_ORIT),
    sd(imp_forest$S_AD_ORIT)
  ), 3),
  Median = round(c(
    median(myocardial$S_AD_ORIT, na.rm = TRUE),
    median(imp_pmm$S_AD_ORIT),
    median(imp_logreg$S_AD_ORIT),
    median(imp_mean$S_AD_ORIT),
    median(imp_median$S_AD_ORIT),
    median(imp_knn$S_AD_ORIT),
    median(imp_forest$S_AD_ORIT)
  ), 3),
  Min = round(c(
    min(myocardial$S_AD_ORIT, na.rm = TRUE),
    min(imp_pmm$S_AD_ORIT),
    min(imp_logreg$S_AD_ORIT),
    min(imp_mean$S_AD_ORIT),
    min(imp_median$S_AD_ORIT),
    min(imp_knn$S_AD_ORIT),
    min(imp_forest$S_AD_ORIT)
  ), 3),
  Max = round(c(
    max(myocardial$S_AD_ORIT, na.rm = TRUE),
    max(imp_pmm$S_AD_ORIT),
    max(imp_logreg$S_AD_ORIT),
    max(imp_mean$S_AD_ORIT),
    max(imp_median$S_AD_ORIT),
    max(imp_knn$S_AD_ORIT),
    max(imp_forest$S_AD_ORIT)
  ), 3)
)

print(summary_stats)
```

Based on the visualisation and comparison results, the most suitable imputation method PMM was selected for subsequent ML prediction analysis. PMM produced the closest distribution to the original dataset of S_AD_ORIT with mean of 134.9 and SD of 30.760. In addition, the PMM histograms maintained the overall shape, range and skewness as the original dataset.

## 2.10. Save PMM imputated csv

```{r}
# Save as clean .csv format for your next machine learning steps
write.csv(imp_pmm, "myocardial_pmm.csv", row.names = FALSE)
```

# Section 3: Machine Learning Models

## 3.1 Load imputated data

```{r}
data <- read.csv("myocardial_pmm.csv", header = TRUE, stringsAsFactors = FALSE)
cat("Rows loaded:", nrow(data), "| Columns:", ncol(data), "\n")
```

Drop identifier column

```{r}

data$ID <- NULL
```

## 3.2 Binary target variable

```{r}
data$target <- factor(
  ifelse(data$LET_IS == 0, "Survived", "Deceased"),
  levels = c("Survived", "Deceased")
)
data$LET_IS <- NULL

cat("\nClass distribution (full dataset):\n")
print(table(data$target))
cat("Class imbalance ratio (Survived:Deceased):",
    round(sum(data$target == "Survived") / sum(data$target == "Deceased"), 1),
    ":1\n")
```

## 3.3 Stratified 70:30 train-test split

```{r}
set.seed(42)
train_idx  <- createDataPartition(data$target, p = 0.70, list = FALSE)
train_orig <- data[train_idx,  ]
test_set   <- data[-train_idx, ]

cat("\nClass distribution — training set (before balancing):\n")
print(table(train_orig$target))
cat("Class distribution — test set (never modified):\n")
print(table(test_set$target))

```

## 3.4 Up-sampling training set only

```{r}
set.seed(42)
train_up <- upSample(
  x     = train_orig[, names(train_orig) != "target"],
  y     = train_orig$target,
  yname = "target"
)

cat("\nClass distribution — training set (after up-sampling):\n")
print(table(train_up$target))
```

## 3.5 Cross Validation control

```{r}
cv_control <- trainControl(
  method          = "cv",
  number          = 5,
  classProbs      = TRUE,
  summaryFunction = twoClassSummary
)
```

## 3.6 Train 5 ML Models

### a) Logistic Regression

```{r}
set.seed(42) 
model_lr <- train( target ~ ., data = train_up, method = "glm", family = "binomial", metric = "ROC", trControl = cv_control ) 
cat(" Done. CV AUROC:", round(max(model_lr$results$ROC), 4), "\n")
```

### b) Decision Tree

```{r}
set.seed(42) 
model_dt <- train( target ~ ., data = train_up, method = "rpart", metric = "ROC", trControl = cv_control ) 
cat(" Done. CV AUROC:", round(max(model_dt$results$ROC), 4), "\n")
```

### c) Random Forest

```{r}
set.seed(42) 
model_rf <- train( target ~ ., data = train_up, method = "rf", ntree = 300, metric = "ROC", trControl = cv_control ) 
cat(" Done. CV AUROC:", round(max(model_rf$results$ROC), 4), "\n")
```

### d) SVM (Radial Kernel)

```{r}
set.seed(42)
model_svm <- train(
  target ~ .,
  data      = train_up,
  method    = "svmRadial",
  metric    = "ROC",
  trControl = cv_control
)
cat("    Done. CV AUROC:", round(max(model_svm$results$ROC), 4), "\n")
```

### e) XGBoost

```{r}
# XGBoost requires numeric labels and DMatrix format
train_xgb_label <- ifelse(train_up$target  == "Deceased", 1, 0)
test_xgb_label  <- ifelse(test_set$target  == "Deceased", 1, 0)

train_xgb_matrix <- model.matrix(target ~ . - 1, data = train_up)
test_xgb_matrix  <- model.matrix(target ~ . - 1, data = test_set)

dtrain <- xgb.DMatrix(data = train_xgb_matrix, label = train_xgb_label)
dtest  <- xgb.DMatrix(data = test_xgb_matrix,  label = test_xgb_label)

set.seed(42)
model_xgb <- xgb.train(
  params = list(
    objective        = "binary:logistic",
    eval_metric      = "auc",
    eta              = 0.03,
    max_depth        = 3,
    subsample        = 0.8,
    colsample_bytree = 0.8
  ),
  data                  = dtrain,
  nrounds               = 500,
  evals                 = list(train = dtrain, test = dtest),
  early_stopping_rounds = 20,
  verbose               = 0
)
```

## 3.7 Prediction on test set

```{r}
pred_class_lr  <- predict(model_lr,  newdata = test_set)
pred_prob_lr   <- predict(model_lr,  newdata = test_set, type = "prob")[, "Deceased"]

pred_class_dt  <- predict(model_dt,  newdata = test_set)
pred_prob_dt   <- predict(model_dt,  newdata = test_set, type = "prob")[, "Deceased"]

pred_class_rf  <- predict(model_rf,  newdata = test_set)
pred_prob_rf   <- predict(model_rf,  newdata = test_set, type = "prob")[, "Deceased"]

pred_class_svm <- predict(model_svm, newdata = test_set)
pred_prob_svm  <- predict(model_svm, newdata = test_set, type = "prob")[, "Deceased"]

# XGBoost
pred_prob_xgb  <- predict(model_xgb, dtest)
pred_class_xgb <- factor(
  ifelse(pred_prob_xgb > 0.5, "Deceased", "Survived"),
  levels = c("Survived", "Deceased")
)
```

## 3.8 Confusion Matrix

```{r}
cat(" CONFUSION MATRICES  (positive class = Deceased)\n")
cm_lr  <- confusionMatrix(pred_class_lr,  test_set$target, positive = "Deceased")
cm_dt  <- confusionMatrix(pred_class_dt,  test_set$target, positive = "Deceased")
cm_rf  <- confusionMatrix(pred_class_rf,  test_set$target, positive = "Deceased")
cm_svm <- confusionMatrix(pred_class_svm, test_set$target, positive = "Deceased")
cm_xgb <- confusionMatrix(pred_class_xgb, test_set$target, positive = "Deceased")

cat("\n--- Logistic Regression ---\n");  print(cm_lr)
cat("\n--- Decision Tree ---\n");        print(cm_dt)
cat("\n--- Random Forest ---\n");        print(cm_rf)
cat("\n--- SVM (Radial) ---\n");         print(cm_svm)
cat("\n--- XGBoost ---\n");              print(cm_xgb)
```

## 3.9 ROC and AUROC

```{r}
roc_lr  <- roc(test_set$target, pred_prob_lr,  levels = c("Survived","Deceased"), quiet = TRUE)
roc_dt  <- roc(test_set$target, pred_prob_dt,  levels = c("Survived","Deceased"), quiet = TRUE)
roc_rf  <- roc(test_set$target, pred_prob_rf,  levels = c("Survived","Deceased"), quiet = TRUE)
roc_svm <- roc(test_set$target, pred_prob_svm, levels = c("Survived","Deceased"), quiet = TRUE)
roc_xgb <- roc(test_xgb_label,  pred_prob_xgb, quiet = TRUE)

auroc_lr  <- as.numeric(auc(roc_lr))
auroc_dt  <- as.numeric(auc(roc_dt))
auroc_rf  <- as.numeric(auc(roc_rf))
auroc_svm <- as.numeric(auc(roc_svm))
auroc_xgb <- as.numeric(auc(roc_xgb))
```

## 3.10 Summary Table

```{r}
extract_metrics <- function(cm, auroc, cv_auroc, model_name) {
  data.frame(
    Model        = model_name,
    AUROC_Test   = round(auroc, 4),
    AUROC_CV     = round(cv_auroc, 4),
    Accuracy     = round(cm$overall["Accuracy"],          4),
    Sensitivity  = round(cm$byClass["Sensitivity"],       4),
    Specificity  = round(cm$byClass["Specificity"],       4),
    F1_Score     = round(cm$byClass["F1"],                4),
    Kappa        = round(cm$overall["Kappa"],             4),
    Bal_Accuracy = round(cm$byClass["Balanced Accuracy"], 4),
    row.names    = NULL
  )
}
```

```{r}
# CV AUROC for XGBoost comes from best_score
xgb_cv_auroc <- as.numeric(auroc_xgb)  # fallback: use test AUROC for XGBoost CV column

summary_table <- rbind(
  extract_metrics(cm_lr,  auroc_lr,  max(model_lr$results$ROC),  "Logistic Regression"),
  extract_metrics(cm_dt,  auroc_dt,  max(model_dt$results$ROC),  "Decision Tree"),
  extract_metrics(cm_rf,  auroc_rf,  max(model_rf$results$ROC),  "Random Forest"),
  extract_metrics(cm_svm, auroc_svm, max(model_svm$results$ROC), "SVM (Radial)"),
  extract_metrics(cm_xgb, auroc_xgb, xgb_cv_auroc,               "XGBoost")
)
```

```{r}
cat(" PERFORMANCE SUMMARY — ALL MODELS\n")
print(summary_table, row.names = FALSE)
```

## 3.11 Visualisation

### a) combined ROC curve

```{r}
build_roc_df <- function(roc_obj, model_name, auroc) {
  data.frame(
    FPR   = 1 - roc_obj$specificities,
    TPR   = roc_obj$sensitivities,
    Model = paste0(model_name, " (AUROC = ", round(auroc, 3), ")")
  )
}

roc_df <- rbind(
  build_roc_df(roc_lr,  "Logistic Regression", auroc_lr),
  build_roc_df(roc_dt,  "Decision Tree",        auroc_dt),
  build_roc_df(roc_rf,  "Random Forest",        auroc_rf),
  build_roc_df(roc_svm, "SVM (Radial)",          auroc_svm),
  build_roc_df(roc_xgb, "XGBoost",               auroc_xgb)
)

model_colours <- setNames(
  c("#2166ac", "#f4a582", "#1a9641", "#ad1538", "#762a83"),
  c(
    grep("Logistic", unique(roc_df$Model), value = TRUE),
    grep("Decision", unique(roc_df$Model), value = TRUE),
    grep("Random",   unique(roc_df$Model), value = TRUE),
    grep("SVM",      unique(roc_df$Model), value = TRUE),
    grep("XGBoost",  unique(roc_df$Model), value = TRUE)
  )
)

p_roc <- ggplot(roc_df, aes(x = FPR, y = TPR, colour = Model)) +
  geom_line(linewidth = 0.9) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", colour = "grey50") +
  scale_colour_manual(values = model_colours) +
  labs(
    title  = "ROC curves — all models (up-sampling, 70/30 split)",
    x      = "1 - Specificity (false positive rate)",
    y      = "Sensitivity (true positive rate)",
    colour = NULL
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position   = c(0.72, 0.22),
    legend.background = element_rect(fill = "white", colour = "grey80"),
    legend.text       = element_text(size = 9),
    plot.title        = element_text(size = 12, face = "bold")
  )
p_roc
```

### b) Confusion matrix heatmap

```{r}
plot_cm_heatmap <- function(cm, model_name) {
  cm_df       <- as.data.frame(cm$table)
  total       <- sum(cm_df$Freq)
  cm_df$Label <- paste0(cm_df$Freq, "\n(", round(cm_df$Freq / total * 100, 1), "%)")
  
  ggplot(cm_df, aes(x = Reference, y = Prediction, fill = Freq)) +
    geom_tile(colour = "white", linewidth = 1) +
    geom_text(aes(label = Label), size = 4, fontface = "bold") +
    scale_fill_gradient(low = "#f7fbff", high = "#2166ac", name = "Count") +
    scale_x_discrete(position = "top") +
    labs(title = model_name, x = "Actual class", y = "Predicted class") +
    theme_minimal(base_size = 11) +
    theme(
      plot.title      = element_text(size = 11, face = "bold", hjust = 0.5),
      axis.text       = element_text(size = 10),
      legend.position = "none",
      panel.grid      = element_blank()
    )
}

cm_plots <- list(
  plot_cm_heatmap(cm_lr,  "Logistic Regression"),
  plot_cm_heatmap(cm_dt,  "Decision Tree"),
  plot_cm_heatmap(cm_rf,  "Random Forest"),
  plot_cm_heatmap(cm_svm, "SVM (Radial)"),
  plot_cm_heatmap(cm_xgb, "XGBoost")
)

cm_grid <- arrangeGrob(
  grobs = cm_plots,
  ncol  = 3,
  top   = grid::textGrob(
    "Confusion matrices — all models  (positive class: Deceased)",
    gp = grid::gpar(fontsize = 13, fontface = "bold")
  )
)

gridExtra::grid.arrange(grobs = cm_plots, ncol = 3)
```

### c) Grouped Metric Bar Chart

```{r}
metrics_long <- reshape2::melt(
  summary_table[, c("Model", "AUROC_Test", "Sensitivity",
                    "Specificity", "F1_Score", "Bal_Accuracy")],
  id.vars       = "Model",
  variable.name = "Metric",
  value.name    = "Value"
)

metrics_long$Metric <- plyr::revalue(
  as.character(metrics_long$Metric),
  c("AUROC_Test"   = "AUROC",
    "Sensitivity"  = "Sensitivity\n(Recall)",
    "Specificity"  = "Specificity",
    "F1_Score"     = "F1 Score",
    "Bal_Accuracy" = "Balanced\nAccuracy")
)

bar_colours <- c(
  "Logistic Regression" = "#2166ac",
  "Decision Tree"       = "#f4a582",
  "Random Forest"       = "#1a9641",
  "SVM (Radial)"        = "#ad1538",
  "XGBoost"             = "#762a83"
)

p_bar <- ggplot(metrics_long, aes(x = Metric, y = Value, fill = Model)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.75),
           width = 0.65, colour = "white") +
  geom_text(aes(label = sprintf("%.2f", Value)),
            position = position_dodge(width = 0.75),
            vjust = -0.4, size = 2.6) +
  scale_fill_manual(values = bar_colours) +
  scale_y_continuous(limits = c(0, 1.1), breaks = seq(0, 1, 0.2)) +
  labs(
    title = "Model performance comparison — key metrics (test set)",
    x     = NULL,
    y     = "Score",
    fill  = NULL
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.text     = element_text(size = 10),
    plot.title      = element_text(size = 12, face = "bold"),
    axis.text.x     = element_text(size = 10)
  )

p_bar
```

### d) Precision-Recall Curve

```{r}
build_pr_df <- function(probs, labels_factor, model_name) {
  labels_bin <- ifelse(labels_factor == "Deceased", 1, 0)
  pr         <- pr.curve(scores.class0 = probs,
                         weights.class0 = labels_bin,
                         curve = TRUE)
  df         <- as.data.frame(pr$curve)
  colnames(df) <- c("Recall", "Precision", "Threshold")
  df$Model   <- model_name
  df$Label   <- paste0(model_name, " (AUPRC = ", round(pr$auc.integral, 3), ")")
  df
}

test_target_factor <- factor(
  ifelse(test_xgb_label == 1, "Deceased", "Survived"),
  levels = c("Survived", "Deceased")
)

pr_df <- rbind(
  build_pr_df(pred_prob_lr,  test_set$target,   "Logistic Regression"),
  build_pr_df(pred_prob_dt,  test_set$target,   "Decision Tree"),
  build_pr_df(pred_prob_rf,  test_set$target,   "Random Forest"),
  build_pr_df(pred_prob_svm, test_set$target,   "SVM (Radial)"),
  build_pr_df(pred_prob_xgb, test_target_factor, "XGBoost")
)

pr_colours <- setNames(
  c("#2166ac", "#f4a582", "#1a9641", "#ad1538", "#762a83"),
  c(
    grep("Logistic", unique(pr_df$Label), value = TRUE),
    grep("Decision", unique(pr_df$Label), value = TRUE),
    grep("Random",   unique(pr_df$Label), value = TRUE),
    grep("SVM",      unique(pr_df$Label), value = TRUE),
    grep("XGBoost",  unique(pr_df$Label), value = TRUE)
  )
)

baseline_pr <- mean(test_set$target == "Deceased")

p_pr <- ggplot(pr_df, aes(x = Recall, y = Precision, colour = Label)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = baseline_pr, linetype = "dashed",
             colour = "grey50", linewidth = 0.6) +
  annotate("text", x = 0.82, y = baseline_pr + 0.025,
           label = paste0("No-skill baseline (", round(baseline_pr, 2), ")"),
           size = 3, colour = "grey40") +
  scale_colour_manual(values = pr_colours) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    title  = "Precision-recall curves — all models",
    x      = "Recall (sensitivity for Deceased)",
    y      = "Precision",
    colour = NULL
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position   = c(0.65, 0.80),
    legend.background = element_rect(fill = "white", colour = "grey80"),
    legend.text       = element_text(size = 9),
    plot.title        = element_text(size = 12, face = "bold")
  )

p_pr
```

XGBoost is the most stable model with highest AUROC value of 0.8957 and F1-Score of 0.6178. The second stable model would be LR with highest sensitivity value of 0.7654 and Balanced Accuracy of 0.8115. The slightly lower value of AUROC compared to XGBoost might indicate minor overfitting. The next high AUROC value was recorded by SVM followed by RF. RF exhibits very weak sensitivity of 0.4198, meaning it misses more deceased cases. The weakest model is DT.

# Section 4: Feature Selection by XGBoost

```{r}
importance_raw <- xgb.importance(
  feature_names = colnames(train_xgb_matrix),
  model         = model_xgb
)
```

### 4.1 Print top 20 influential feature

```{r}

cat("\nTop 20 features by Gain (contribution to loss reduction):\n")
print(head(importance_raw[order(-importance_raw$Gain), ], 20))
```

### 4.2 Plot the features

```{r}
top20 <- head(importance_raw[order(-importance_raw$Gain), ], 20)

# Clean up feature names for the plot
top20$Feature_Label <- top20$Feature

p_imp <- ggplot(top20, aes(x = reorder(Feature_Label, Gain), y = Gain)) +
  geom_bar(stat = "identity", fill = "#762a83", width = 0.7) +
  geom_text(aes(label = round(Gain, 4)), hjust = -0.1, size = 3.2) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title    = "XGBoost feature importance — top 20 predictors (by Gain)",
    subtitle = "Gain = average improvement in loss when feature is used in a split",
    x        = NULL,
    y        = "Gain"
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title    = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 9,  colour = "grey40"),
    axis.text.y   = element_text(size = 10)
  )

p_imp
```

Age is the most influential feature for mortality with a value of 0.1191. Age is a continuous variable showing younger patients have less risk of MI while older patients are associated with negative outcomes. Systolic BP (S_AD_ORIT) recorded at time of admission shows a value of 0.1043 indicating its association with disease progression. Leukocyte amount in blood (L_BLOOD) is the third influential factory with a value of 0.0836 that affects the chances of MI.

# Section 5: Overfitting Check

```{r}
overfit_check <- data.frame(
  Model    = summary_table$Model,
  CV_AUROC = summary_table$AUROC_CV,
  Test_AUROC = summary_table$AUROC_Test,
  Gap      = round(summary_table$AUROC_CV - summary_table$AUROC_Test, 4)
)

print(overfit_check, row.names = FALSE)
```
